---
title: "Ideological Innocence, Revisited"
subtitle: "A Computational Reproduction and New-Data Replication of Kalmoe (2020) — FORRT ReplicaTHIS Nomination #8"
author: "Lukas Wallrich"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    code-tools: true
    embed-resources: true
    theme: cosmo
    fig-width: 8
    fig-height: 5
    tbl-cap-location: top
execute:
  echo: true
  warning: false
  message: false
  cache: false
bibliography: references.bib
---

```{r}
#| label: setup
#| include: false

# Anchor here() to the project root regardless of render location.
here::i_am("manuscript/manuscript.qmd")
library(here)
library(dplyr)
library(tidyr)
library(gt)
library(ggplot2)

# ---- Load all precomputed outputs (the qmd only READS these) ----------------
# Reproduction (Tables 1,2,4,5, Fig 1, footnote 5, in-text)
R_t1     <- readRDS(here("output/repro_main/table1.rds"))
R_t2     <- readRDS(here("output/repro_main/table2.rds"))
R_t4     <- readRDS(here("output/repro_main/table4.rds"))
R_t5     <- readRDS(here("output/repro_main/table5.rds"))
R_intext <- readRDS(here("output/repro_main/intext_stats.rds"))
R_fn5    <- readRDS(here("output/repro_main/footnote5.rds"))
R_fig1   <- readRDS(here("output/repro_main/fig1_data.rds"))

# Panels (Table 3)
P        <- readRDS(here("output/repro_panels/table3.rds"))

# Robustness
RC1 <- readRDS(here("output/robustness/rc1_polar.rds"))
RC2 <- readRDS(here("output/robustness/rc2_nonattitude.rds"))
RC4 <- readRDS(here("output/robustness/rc4_stratifier.rds"))
RC6 <- readRDS(here("output/robustness/rc6_vote.rds"))
RC7 <- readRDS(here("output/robustness/rc7_time.rds"))

# Replication (ANES 2020/2024 + panel)
X_t1   <- readRDS(here("output/replication/table1_breadth.rds"))
X_t2   <- readRDS(here("output/replication/table2_reliability.rds"))
X_t4   <- readRDS(here("output/replication/table4_correlations.rds"))
X_t5   <- readRDS(here("output/replication/table5_vote.rds"))
X_stab <- readRDS(here("output/replication/table3_stability.rds"))
X_fig1 <- readRDS(here("output/replication/fig1_distributions.rds"))
X_opin <- readRDS(here("output/replication/opinionation.rds"))
X_cmp  <- readRDS(here("output/replication/comparison_fullsample.rds"))
X_kn   <- readRDS(here("output/replication/knowledge.rds"))

# GSS
G1 <- readRDS(here("output/gss/analysis1_polar.rds"))
G2 <- readRDS(here("output/gss/analysis2_reliability.rds"))
G3 <- readRDS(here("output/gss/analysis3_correlations.rds"))
G4 <- readRDS(here("output/gss/analysis4_probit.rds"))
G5 <- readRDS(here("output/gss/analysis5_timetrend.rds"))
G6 <- readRDS(here("output/gss/analysis6_polviews_dk.rds"))

# Revision analyses (R/07_revision.R), added in response to the peer review
V_A <- readRDS(here("output/revision/A_sorting_constraint.rds"))     # sorting vs constraint
V_B <- readRDS(here("output/revision/B_ame_ratios.rds"))             # AME-based H÷L ratios
V_C <- readRDS(here("output/revision/C_panel_attrition.rds"))        # panel attrition/reweighting
V_D <- readRDS(here("output/revision/D_breadth_nonattitudes.rds"))   # breadth with/without nonattitudes
V_E <- readRDS(here("output/revision/E_mode_comparison.rds"))        # 2020 mode comparison
V_F <- readRDS(here("output/revision/F_small_computations.rds"))     # small computations
V_G <- readRDS(here("output/revision/G_bootstrap_ratio.rds"))        # bootstrap CI for 2020 ratio

# ---- Formatting + accessor helpers ------------------------------------------
d1 <- function(x) sprintf("%.1f", x)
d2 <- function(x) sprintf("%.2f", x)
pp <- function(x) sprintf("%.0f", x)          # percentage, no decimals

# Reproduction accessors
t1v <- function(con, col) R_t1[[col]][R_t1$construct == con]
t2v <- function(con, st, col) R_t2[[col]][R_t2$construct == con & R_t2$stat == st]
t4v <- function(pr, col) R_t4[[col]][R_t4$pair == pr]
t5v <- function(pr, st, col) R_t5[[col]][R_t5$predictor == pr & R_t5$stat == st]
ix  <- function(nm) { s <- R_intext$summary; s$reproduced[s$stat == nm] }

# Replication accessors
x5 <- function(tbl, pr, st, col) { d <- tbl; d[[col]][d$predictor == pr & d$stat == st] }

# Revision accessors
va <- function(era, grp, col) V_A[[col]][V_A$era == era & V_A$group == grp]   # Module A cell
vb <- function(era, pr, col) V_B[[col]][V_B$era == era & V_B$predictor == pr] # Module B cell
```

## Abstract {.unnumbered}

Kalmoe's [-@kalmoe2020] *Uses and Abuses of Ideology in Political Psychology* argues that
polar, coherent, stable, and electorally potent ideology characterises only the most
politically knowledgeable minority of Americans, while partisanship is strong for everyone.
Responding to a
FORRT ReplicaTHIS nomination [@forrt_nomination], I computationally reproduce the article's
public-data analyses, subject them to eight robustness checks, and replicate them in data the
author never saw: ANES 2020 and 2024 and the 2020&rarr;2024 panel, with a General Social Survey
(GSS) cross-dataset test. The reproduction is faithful: of 176 table cells, 154 match tightly and
only five diverge, alongside three documentation-level anomalies. Every substantive conclusion
survives the robustness suite, though one check reframes the "breadth" statistic as largely an
expression of how nonattitudes are coded. The replication is two-sided. Mass ideological coherence,
polarity, stability, and bivariate vote potency all rose sharply, so the "innocence" portrait is
time-bound as stated. The knowledge gradient then splits by metric: ideological *coherence* stays
stratified (the low-knowledge floor rose, but a gradient remains), whereas *vote potency*
compressed sharply. A measure-matched ideology gradient of roughly 10.5&times; in 1988&ndash;92
falls to about 2.5&times; in 2020 and 2.0&times; in 2024 on pseudo-R², and on the
average-marginal-effect scale it was already &le;2&times; and now falls below one. However, a
party-controlled test qualifies this: net of party identification, ideology's *independent* vote
contribution did not rise overall or at the low-knowledge floor, so the bivariate rise largely
reflects ideology's growing alignment with party rather than a new stratum of independent
constraint. Partisanship still dominates every metric, though the gap narrowed. The GSS reproduces
the depth gradients almost exactly while the breadth gradient does not, an instructive dissociation.

## Introduction

Since Converse's [-@converse1964] account of belief systems in mass publics, political science
has debated whether ordinary citizens organise their politics ideologically. Kalmoe's [-@kalmoe2020]
contribution, drawing on the framework of *Neither Liberal nor Conservative* [@kinderkalmoe2017],
makes a sustained empirical case that they largely do not. Using the ANES Time-Series Cumulative Data
File and three ANES panels, Kalmoe shows that self-identified ideology and core political values are
broad, coherent, stable, and predictive of the vote only among the most knowledgeable citizens. The
stratification he documents is enormous, with highest-to-lowest "info-gain" ratios that range from
about 1.6&times; to more than 10&times;. Party identification, by contrast, is strong and
knowledge-robust across the board. The paper closes with four "best practices" for political psychology: stratify by knowledge,
do not generalise from high-knowledge convenience samples, distinguish ideology from partisanship,
and treat full-sample estimates as knowledge-conditioned.

The FORRT ReplicaTHIS nomination [@forrt_nomination] asks whether these conclusions are sensitive
to operationalisation, time period, issue selection, and how the knowledge moderator is measured.
I address that question with three complementary exercises, each of which answers a different
question. A **computational reproduction** asks whether the published numbers can be regenerated
from the same public data and stated methods, and so tests the paper's internal integrity. A
**robustness analysis** asks whether the conclusions survive defensible alternative choices, and so
tests their fragility. A **new-data replication** asks whether the findings recur in data the author
never saw, and so tests their generality and their durability in a changed political era. The
original had no code or data package [see Methods], so I had to reconstruct the pipeline from the
paper and its online supplement, adjudicating unstated choices against the anchors the tables
themselves provide.

## Data and Method

### Data provenance

The reproduction uses the **ANES Time-Series Cumulative Data File**, 1948&ndash;2016 vintage
(59,944 &times; 1,029), redistributed as an unmodified `.rda` by the `anesr` package. This is the
release available when Kalmoe submitted the paper, and is therefore a *better* basis for a faithful
reproduction than a current release, whose recodes could differ [@anes_cdf]. The three panels
(1990&ndash;92, 1992&ndash;96, 2000&ndash;02) come from the corresponding `anesr` year files. The
replication uses ANES **2020** ($n = 8{,}280$) and **2024** ($n = 5{,}521$) in cumulative-file
coding, obtained as a custom subset of the 1948&ndash;2024 ANES Cumulative Data File from SDA
Berkeley (accessed 2026-07-03; no login), supplemented by the 2024 knowledge battery and the
2016&ndash;2020&ndash;2024 panel link from the same source [@anes2020; @anes2024]. The GSS test
uses the **GSS 1972&ndash;2024 Cumulative Datafile** (Release 3, 75,699 respondents), also a
public SDA Berkeley subset [@gss]. The 1984&ndash;2016 rows appear in both the `anesr` and the SDA
files, allowing a cross-source integrity check.

### The knowledge moderator and the anchor-based strategy

The paper turns on a single moderator: a **5-point interviewer rating** of each respondent's
general political information, coded 0/.25/.5/.75/1. It is worth noting that the five knowledge
groups are the five *fixed rating categories*, whose natural population shares are uneven (roughly
9/20/34/25/13), not quantiles. Because the original documents no variable codes, item sets,
weights, or edge-case rules, each unstated choice was adjudicated **empirically against the anchors
the tables supply**: the construct sample sizes, the published group shares, and the printed
polar-half percentages. This is not tuning; the anchors identify the author's actual choices. For
example, the pre-election interviewer rating (`VCF0050a`), with the post-election rating filling
the four midterm gaps, reproduces Table 1's ideology row *exactly* by group and minimises total
error across all 30 cells, confirming both the variable and the "pre-preferred" combination.

### Key judgment calls

The consequential choices, condensed from the decisions log and reports, were: (1) the pre-preferred
combined interviewer rating as moderator, grouped on the five fixed levels; (2) restriction to
face-to-face and telephone interviews (excluding the 2012/2016 web samples), matching the construct
$N$s within about 4%; (3) per-table weighting &mdash; Tables 1 and 5 weighted (`VCF0009x`, the
FTF-sample weight), Figure 1 and Tables 2 and 4 unweighted, exactly as the paper states; (4) the
polar-half boundary as **inclusive** ($|score| \geq 0.5$), which alone reproduces the value-index
percentages; (5) ideology "haven't thought much about it" (HTMA) coded with moderates at 0, and
policy "don't know" coded to the scale midpoint &mdash; the latter independently confirmed by
Table 2 reliability, not just Table 1; and (6) indices computed as the mean of available items.
Construct codes were resolved to the specific cumulative-file variables that reproduce the anchors
(e.g. moral traditionalism is the 5-point agree&ndash;disagree battery `VCF0851`&ndash;`VCF0854`,
*not* the government-support dichotomies). The replication mirrors this coding exactly, deviating
only where 2020/2024 data force it (see §5).

### Independent verification

Before any number entered this manuscript, every headline result from all five pipelines was
**independently recomputed** by a separate verification pass that loaded the raw data directly and
hand-rolled its own coders rather than re-running the analysis functions. The verification confirmed
the reproduction cell-by-cell, the panel merges (100% clean joins), the replication construct-coding
consistency, the stability panel join, and the GSS gradients. It found **one genuine bug**: a
mislabelled ratio in the replication's vote-stratification headline, where a column computed as
Full&divide;Lowest was labelled Highest&divide;Lowest, producing a spurious "10.5&rarr;1.7" collapse.
The corrected, measure-matched Highest&divide;Lowest values (about 2.5 in 2020, 2.0 in 2024) are used
throughout below; the fix is reported transparently and is documented in the verification record.
Two further items were relabelling cautions (pairwise versus average correlations; "collapse" versus
"halving" language), both incorporated here. The verification scripts live in `output/verification/`
(`v_reproduction.R`, `v_robustness.R`, `v_panels.R`, `v_replication.R`, `v_gss.R`) and are listed in
the Reproducibility Statement. All five pipelines regenerate byte-identically.

## Reproduction Results

### Fidelity

```{r}
#| label: tbl-fidelity
#| tbl-cap: "Reproduction fidelity, banded by absolute cell difference. Cross-sectional tables (1, 2, 4, 5) use the per-table tolerances of §3 (Table 1 ≤1/≤2 pp; Tables 2/4 ≤.01/≤.02; Table 5 coef ≤.05/≤.10, R² ≤.01/≤.02). Panel stability (Table 3) uses fixed r² bands (exact ≤.005 / within .01 / within .02 / beyond), computed from output/repro_panels/comparison.csv."
pb <- V_F$panel_fidelity_band
prow <- function(p) as.integer(pb[pb$panel == p, c("exact_le005","within_01","within_02","beyond_02")])
fidelity <- tibble::tibble(
  Component = c("Tables 1, 2, 4, 5 (176 cells)",
                "Table 3 — 1990–92 (25 cells)",
                "Table 3 — 1992–96 (25 cells)",
                "Table 3 — 2000–02 (15 cells)"),
  `Exact / Tight` = c(154L, prow("1990-92")[1], prow("1992-96")[1], prow("2000-02")[1]),
  `Within .01 / Close` = c(17L, prow("1990-92")[2], prow("1992-96")[2], prow("2000-02")[2]),
  `Within .02` = c(NA_integer_, prow("1990-92")[3], prow("1992-96")[3], prow("2000-02")[3]),
  `Beyond / Diverge` = c(5L, prow("1990-92")[4], prow("1992-96")[4], prow("2000-02")[4])
)
fidelity |>
  gt() |>
  sub_missing(columns = everything(), missing_text = "—") |>
  tab_header(title = "Cross-sectional reproduction is faithful; panel fidelity is more mixed",
             subtitle = "Every row sums to its cell total") |>
  cols_align("left", columns = Component)
```

Across the four cross-sectional tables, 154 of 176 cells match tightly, 17 are close, and five
diverge (@tbl-fidelity). Tables 2 and 4 reproduce essentially exactly, with maximum absolute
differences of only .015 and .018. Four of the five diverges are Table 1 *Lowest-knowledge* cells,
the smallest and noisiest stratum, which comes out at about 7% of the sample here against the paper's
9%; the fifth is one Table 5 coefficient in the small *Lower* voter group. The panels reproduce less
evenly. Policy Views and Ideology ID come within &plusmn;.01 in every cell of all three panels, the
one documented exception being the 2000&ndash;02 policy row (§5.3). The value batteries,
egalitarianism and moral traditionalism, reproduce less cleanly in the 1992&ndash;96 and
2000&ndash;02 panels, where several cells fall beyond .02. The banding in @tbl-fidelity is exclusive,
and every row sums to its stated cell count.

**Anchor cells versus validation cells.** Because unstated coding choices were adjudicated against
the anchors the tables supply (§2.2), cells used to *fix* a coding decision cannot also *validate*
it. I therefore partition the 176 cross-sectional cells into two sets. The anchor cells are the
Table 1 full-sample and ideology-row cells, which fixed the pre/post rating, the weighting, and the
polar-half boundary, together with the Table 2 policy-&alpha; cells, which fixed the DK&rarr;midpoint
rule. The remaining cells form an independent-validation set. The anchor set comprises
`r V_F$anchor_partition$n[V_F$anchor_partition$role=="anchor"]` cells, with a tight-match rate of
`r d1(V_F$anchor_partition$tight_pct[V_F$anchor_partition$role=="anchor"])`%. The
`r V_F$anchor_partition$n[V_F$anchor_partition$role=="validation"]` validation cells, which include
all of Table 4 and most of Tables 2 and 5 and were matched *without* being used to set any coding
choice, still match tightly at
`r d1(V_F$anchor_partition$tight_pct[V_F$anchor_partition$role=="validation"])`%, and all five
diverges fall in this set. Even under the most generous anchoring, treating *all* 30 Table 1 cells
plus the policy-&alpha; row as anchors
(`r V_F$anchor_partition_generous$n[V_F$anchor_partition_generous$role=="anchor"]` cells), the
validation tight-match rate is
`r d1(V_F$anchor_partition_generous$tight_pct[V_F$anchor_partition_generous$role=="validation"])`%.
The agreement is therefore genuinely informative rather than mechanical: the cells that did no
adjudication work reproduce as faithfully as those that did.

### Cross-sectional tables

For breadth (Table 1), the reproduced full-sample polar-half shares match the published values
closely: egalitarianism `r pp(t1v("egal","Full"))`% (published 32), moral traditionalism
`r pp(t1v("mt","Full"))`% (35), policy views `r pp(t1v("policy","Full"))`% (18), ideology
`r pp(t1v("ideo","Full"))`% (27), and partisanship `r pp(t1v("party","Full"))`% (61). The ideology
row reproduces *exactly* by knowledge group, and the broad info-gain ordering is preserved, with
ideology remaining steepest (H&divide;L `r d2(t1v("ideo","H_div_L"))`). Two pairs do swap rank,
however. Egalitarianism comes out marginally flatter here (`r d2(t1v("egal","H_div_L"))`) than
partisanship (`r d2(t1v("party","H_div_L"))`), against the original's 1.65 versus 1.63, and the
middle pair swaps too, with policy at `r d2(t1v("policy","H_div_L"))` against moral traditionalism at
`r d2(t1v("mt","H_div_L"))`, versus the original's 2.64 and 2.65. These are hairline reorderings
within a reproduced gradient, not substantive divergences (@tbl-repro-headline).

```{r}
#| label: tbl-repro-headline
#| tbl-cap: "Reproduction versus original for the headline breadth statistic (Table 1) and vote potency (Table 5, full-sample pseudo-R²). Original values transcribed from the published tables."
orig_t1_full  <- c(egal=32, mt=35, policy=18, ideo=27, party=61)
orig_t1_hl    <- c(egal=1.65, mt=2.65, policy=2.64, ideo=4.00, party=1.63)
orig_t5_r2    <- c(Egalitarianism=.15, `Moral Tradition`=.13, `Policy Views`=.21,
                   `Ideology ID`=.23, Partisanship=.49)
cons <- c("egal","mt","policy","ideo","party")
labs <- c("Egalitarianism","Moral Traditionalism","Policy Views","Ideology ID","Partisanship")
t5map <- c("Egalitarianism","MoralTrad","PolicyViews","IdeoID","Partisanship")

headline <- tibble::tibble(
  Construct = labs,
  `Polar-half %, orig` = orig_t1_full[cons],
  `Polar-half %, repro` = round(sapply(cons, function(c) t1v(c, "Full"))),
  `Info-gain H÷L, orig` = orig_t1_hl[cons],
  `Info-gain H÷L, repro` = round(sapply(cons, function(c) t1v(c, "H_div_L")), 2),
  `Vote R², orig` = orig_t5_r2,
  `Vote R², repro` = round(sapply(t5map, function(p) t5v(p, "pseudoR2", "Full")), 2)
)
headline |>
  gt() |>
  tab_header(title = "Original vs reproduced: breadth and vote potency",
             subtitle = "Full-sample values; per-cell grids in the appendix") |>
  cols_align("left", columns = Construct) |>
  tab_spanner("Breadth (Table 1)", columns = 2:5) |>
  tab_spanner("Vote potency (Table 5)", columns = 6:7)
```

Reliability (Table 2) reproduces the steep coherence gradient: egalitarianism &alpha; rises from
`r d2(t2v("egal","alpha","Lowest"))` in the lowest group to `r d2(t2v("egal","alpha","Highest"))`
in the highest, and policy from `r d2(t2v("policy","alpha","Lowest"))` to
`r d2(t2v("policy","alpha","Highest"))`, matching the published .38&rarr;.80. The vote models
(Table 5) reproduce partisanship's dominance, with a full-sample pseudo-R² of
`r d2(t5v("Partisanship","pseudoR2","Full"))` against ideology's
`r d2(t5v("IdeoID","pseudoR2","Full"))`, as well as the sharp within-construct gradient, ideology
rising from `r d2(t5v("IdeoID","pseudoR2","Lower"))` to `r d2(t5v("IdeoID","pseudoR2","Highest"))`
across voter knowledge groups. Figure 1's qualitative story reproduces as well: the ideology
distribution is dominated by a Moderate/HTMA bar of
`r d1(R_fig1$ideology$pct[R_fig1$ideology$ideo_cat==4])`% (paper &asymp;49%), while partisans populate
the poles. Footnote 5's abortion probit reproduces the "roughly twice / about five times" contrast,
with a coefficient ratio of `r d1(R_fn5$coef_ratio[1])`&times; and a pseudo-R² ratio of
`r d1(R_fn5$r2_ratio[1])`&times;.

The reproducible in-text statistics also come out close. Mean knowledge is
`r d2(ix("knowledge_HTMA_ideology"))` for HTMA identifiers against
`r d2(ix("knowledge_libcon_identifiers"))` for liberals and conservatives (paper .36 / .63);
`r pp(ix("policy_pct_answered_all_five")*100)`% answer all five policy items (66%); and the
individual-level ideology&times;vote correlation is `r d2(ix("r_ideology_vote"))` (paper .49). That
last figure sits against the .90 that Jost [-@jost2006] reported from *aggregated* data, and it is
precisely the disaggregation gap that motivates the paper.

<details>
<summary>Full cell-level reproduction grids (Tables 1, 2, 4, 5)</summary>

```{r}
#| label: tbl-repro-full
#| tbl-cap: "Reproduced Table 1 (breadth, %) by knowledge group."
R_t1 |>
  transmute(Construct = labs[match(construct, cons)], N,
            Full = round(Full,1), Lowest = round(Lowest,1), Low = round(Low,1),
            Middle = round(Middle,1), High = round(High,1), Highest = round(Highest,1),
            `H-L` = round(H_minus_L,1), `H÷L` = round(H_div_L,2)) |>
  gt() |> tab_header(title = "Table 1 (reproduced): percent in polar half")
```

```{r}
#| label: tbl-repro-t4
#| tbl-cap: "Reproduced Table 4 (Pearson correlations among constructs) by knowledge group."
R_t4 |>
  mutate(across(Full:Highest, ~round(.x,2))) |>
  select(-N) |>
  gt() |> tab_header(title = "Table 4 (reproduced): construct correlations")
```

```{r}
#| label: tbl-repro-t5
#| tbl-cap: "Reproduced Table 5 (presidential-vote probit) by voter knowledge group."
R_t5 |>
  mutate(across(Full:Highest, ~round(.x,2))) |>
  gt() |> tab_header(title = "Table 5 (reproduced): vote probit coefficient / SE / pseudo-R²")
```

</details>

### Three documentation-level anomalies

Three numbers resist reproduction. I read them as minor errata or documentation gaps in an
otherwise fully reproducible paper.

**(i) A probable N transposition in Table 1.** The reproduced construct sample sizes are ideology
$N = `r format(t1v("ideo","N"), big.mark=",")`$ and partisanship
$N = `r format(t1v("party","N"), big.mark=",")`$, with party the larger of the two, as it should be,
since party ID has near-zero item nonresponse while the ideology item loses non-placers. The paper
prints the reverse, with ideology at 25,332 above partisanship at 24,307. The crosswise match, my
party against the paper's 25,332 and my ideology against its 24,307, is far tighter than the printed
assignment, so the two rows' $N$s were most likely transposed in the original. I flag this as an
erratum candidate rather than silently "correcting" it.

**(ii) The 1990&ndash;92 egalitarianism Middle cell.** Every other egalitarianism cell in that panel
reproduces exactly (.24 / .14 / &mdash; / .32 / .38), as do all the moral-traditionalism, policy, and
ideology cells on the identical subsample and grouping. Yet the Middle group's squared stability comes
out at .14 against the published .26, and no tested variant reaches .26, whether alternative DK
handling, complete-case indices, or leave-one-out over the six items. I cannot reproduce this single
cell, and flag it as a likely erratum.

**(iii) The 2000&ndash;02 policy stability row.** The 2002 study carries none of the five classic
7-point policy scales, so this row cannot be reconstructed from any documented item set. The closest
principled reconstruction, nine federal-spending items common to both waves, matches the published
$N$ almost exactly (1,017 versus 1,016) but yields higher stability (.39 against the published .27).
I report this row as an approximation rather than a validated reproduction, since the item set Kalmoe
used is undocumented in the paper and its supplement.

## Robustness

Eight checks probe exactly the sensitivities the nomination raises, namely operationalisation, time
period, issue selection, and measurement of the moderator, and each reuses the reproduction's coding
so that cells stay comparable. Where the author's online supplement provides a target (Tables A1,
B1&ndash;B5, C1&ndash;C5, D4), I use it as an external validation anchor. @tbl-robustness summarises
the verdicts.

```{r}
#| label: tbl-robustness
#| tbl-cap: "Robustness suite: eight checks, each validated against the author's supplement where available."
rob <- tibble::tibble(
  Check = c("RC1 Polar-half definition", "RC2 Nonattitude coding",
            "RC3 Weighting", "RC4 Stratifier alternatives",
            "RC5 Reliability (ordinal α, ω)", "RC6 Vote models (multivariate / link / AME)",
            "RC7 Time periods", "RC8 Issue selection"),
  Probes = c("exclusive / ±1 SD / outer-third boundary",
             "HTMA & DK → drop vs → midpoint",
             "weight Tables 2 & 4",
             "education / quiz / median split",
             "polychoric α and McDonald's ω",
             "values beyond party+ideo; logit/LPM; marginal effects",
             "1984–96 vs 2000–16; 2008–16",
             "leave-one-out + 6-item index"),
  Verdict = c("Survives", "Qualifies framing", "Survives", "Survives",
              "Survives", "Survives", "Survives", "Survives")
)
rob |>
  gt() |>
  tab_header(title = "Every substantive conclusion survives",
             subtitle = "RC2 reframes the breadth statistic without overturning it") |>
  cols_align("left", columns = everything()) |>
  data_color(columns = Verdict,
             fn = function(x) ifelse(x == "Survives", "#e7f0e9", "#fdf3e2"))
```

The two checks worth dwelling on are RC2 and RC6c.

**RC2: the breadth gradient largely expresses nonattitude coding.** The paper scores ideology
HTMA and policy DK at the scale midpoint. When these nonattitudes are instead *dropped*, the Table 1
breadth gradient flattens. The highest-minus-lowest gap is **eliminated for policy** (from +17 to
about +1.6 points) and **halved for ideology** (from +33 to about +17), because the lowest-knowledge
group jumps once it is restricted to respondents who actually hold an attitude. HTMA is
`r d1(RC2$htma_share_pct)`% of the ideology sample, and the famous Figure 1 middle bar is about half
nonattitude, so once HTMA is excluded the true-moderate bar is
`r d1(RC2$moderate_bar$dropped_true_moderate_pct)`% rather than 50%. This does not overturn the paper;
instead it operationalises its "innocence" mechanism, in which low-knowledge citizens lack attitudes
to place, a selection effect rather than a coding trick. However, it does mean that the polar-half
statistic should not be read as evidence independent of nonresponse. The *coherence* gradient (Table 4
correlations) is more robust to the coding, since dropping nonattitudes leaves the ideology&times;party
gradient similar, at `r d2(RC2$cor_ideo_drop$Lowest[RC2$cor_ideo_drop$pair=="ideo x party"])` in the
lowest group and `r d2(RC2$cor_ideo_drop$Highest[RC2$cor_ideo_drop$pair=="ideo x party"])` in the
highest. This comparison is not strictly like-for-like, though. Because HTMA is concentrated in the
lowest-knowledge group, dropping it removes a large share of that group but little of the top, so the
surviving low-knowledge respondents are the more opinionated subset. The gradient among
attitude-holders is genuine, but it is computed on a differentially selected sample, driven by the
same selection mechanism that produces the breadth result, and so it does not prove that coherence is
untouched by how nonattitudes are handled.

**RC6c: potency stratification on the effect-size scale.** The vote-choice stratification looks
enormous as a variance-explained ratio, with a pseudo-R² highest&divide;lowest of 5&ndash;10&times;,
but on the interpretable probability scale it is moderate. Average marginal effects (AMEs) give a
highest&divide;lowest ratio of about 1.4&ndash;1.7&times; across constructs, just below the
probit-coefficient ratio of about 1.9&ndash;2.7&times;, and far below the pseudo-R² ratio of
5&ndash;10&times;. The gap is not a simple squaring. McFadden's pseudo-R² is a log-likelihood-ratio
statistic ($1 - \mathrm{LL}/\mathrm{LL}_0$) rather than a variance-explained proportion, and in a
probit model the translation from the coefficient to fit depends on the base rate, the within-group
predictor variance, and the curvature of the link. The pseudo-R² ratio therefore inflates group
differences relative to the interpretable probability scale by an amount that varies with those
quantities, rather than tracking any fixed power of the coefficient ratio. One result stands out:
**partisanship's marginal effect on vote probability is flat, indeed slightly declining, across
knowledge** (AME ratio `r d2(RC6$ame$AME_H_div_L[RC6$ame$predictor=="party"])`), even as its
coefficient rises. On the scale that matters for behaviour, partisanship is not merely the strongest
predictor but the most *uniformly* strong, which reinforces the paper's knowledge-robustness claim.

The remaining checks confirm cleanly. The breadth gradient is invariant to the polar-half boundary
(RC1); weighting moves no cell by more than about .02 (RC3); education and a campaign-knowledge quiz
reproduce the gradient (validating against supplement Tables B1&ndash;B5), while a binary median split
hides roughly half of it (RC4); ordinal reliability leaves the gradient if anything steeper (RC5,
validated against Table D4); adding the two values to a party+ideology vote model raises pseudo-R² by
only about 2&ndash;4 points (RC6a); no single policy item is load-bearing (RC8); and coherence,
together with its stratification, is if anything *stronger* in the polarised 2000&ndash;2016
sub-period than in 1984&ndash;1996 (RC7, validated against Tables C1&ndash;C5). That last result
foreshadows the replication.

## Replication in ANES 2020 and 2024

The original used data through 2016. Whether its conclusions hold in the hyper-polarised 2020s is a
genuine test of generality, and one the author could not run.

### Forced deviations

Four deviations are unavoidable, and each is handled conservatively. **(1) Knowledge measure.** The
interviewer rating was discontinued, and is empty for every 2020/2024 case, so knowledge is a
proportion-correct **civics quiz**, precisely the alternative the author himself uses in supplement
Tables B1&ndash;B5. The four pre-election items take five values that provide the five knowledge groups
directly. Their achieved population shares (2020: `r paste(X_kn$shares_2020$pre_only, collapse="/")`;
2024: `r paste(X_kn$shares_2024$pre_only, collapse="/")`, lowest to highest) approximate Kalmoe's fixed
interviewer-rating shares of 9/20/34/25/13 rather than being re-quantiled, so the lowest group is
comparably populated across eras (about 10% in 2020 and `r d1(X_kn$shares_2024$pre_only[1])`% in 2024),
and the merged voter groups behind the vote tables carry the cell $n$s shown there. Because the
author's own supplement shows quiz stratification is "slightly less distinct" than the interviewer
rating, every stratification comparison below is anchored to his *quiz-stratified* original-era tables
(quiz-versus-quiz), not to the steeper interviewer-rated tables. **(2) Web mode.** 2020 is 94% web and
2024 about 81% web, so the FTF/phone restriction cannot be applied. I therefore compare the key
ideology metrics across mode within 2020 (§ "The shrinking middle", @tbl-rep-mode): the HTMA rate is
mode-independent, but the non-web cells are too small to settle whether the moderate-to-polar shift is
mode-driven. **(3) Reduced value batteries.** The cumulative file carries only 4 of 6 egalitarianism
items and 2 of 4 moral-traditionalism items for these years. Since fewer items mechanically inflate
polar-half shares and depress &alpha;, the value comparisons use a **matched-item 1984&ndash;2016
baseline** and lead with item-count-invariant covariance. **(4) Vote.** The two-party recode
`VCF0704a` is empty for 2024, so vote is derived from `VCF0704`.

### Cross-sectional results

Every metric of mass ideology rose. Breadth, coherence, correlation, and vote-predictiveness all
increased relative to both the 1984&ndash;2016 and the 2008&ndash;2016 anchors (@tbl-rep-crosssec).

```{r}
#| label: tbl-rep-crosssec
#| tbl-cap: "Full-sample analogues of Tables 1, 4, and 5 across eras. Egalitarianism/moral-traditionalism breadth is shown against its matched-item 1984–2016 baseline in brackets. Ideology & partisanship are on unchanged scales."
af <- X_cmp$anchor_full
matched <- X_cmp$matched_baseline
# breadth with matched brackets for egal/mt
br8416 <- af$breadth_8416
br_lab <- ifelse(af$construct %in% c("Egalitarianism","Moral Tradition"),
                 paste0(br8416, " [", matched$breadth_matched_8416[match(af$construct, matched$construct)], "]"),
                 as.character(br8416))
rep_cs <- tibble::tibble(
  Construct = af$construct,
  `Breadth 1984–2016` = br_lab,
  `Breadth 2008–16` = round(af$breadth_0816),
  `Breadth 2020` = round(X_cmp$breadth_2020),
  `Breadth 2024` = round(X_cmp$breadth_2024),
  `Avg corr 1984–2016` = af$avgcorr_8416,
  `Avg corr 2008–16` = round(af$avgcorr_0816, 2),
  `Avg corr 2020` = round(X_cmp$avgcorr_2020, 2),
  `Avg corr 2024` = round(X_cmp$avgcorr_2024, 2),
  `Vote R² 1984–2016` = af$voter2_8416,
  `Vote R² 2008–16` = round(af$voter2_0816, 2),
  `Vote R² 2020` = round(X_cmp$voteR2_2020, 2),
  `Vote R² 2024` = round(X_cmp$voteR2_2024, 2)
)
rep_cs |>
  gt() |>
  tab_header(title = "Mass ideology rose on every metric, 2016 → 2020/2024",
             subtitle = "Breadth = % polar half; correlation = mean inter-construct r; vote = pseudo-R²") |>
  tab_spanner("Breadth (%)", columns = 2:5) |>
  tab_spanner("Avg. correlation", columns = 6:9) |>
  tab_spanner("Vote pseudo-R²", columns = 10:13) |>
  cols_align("left", columns = Construct) |>
  fmt_markdown(columns = `Breadth 1984–2016`) |>
  tab_options(table.font.size = pct(85), container.overflow.x = "auto")
```

Ideology's polar share reached `r pp(X_cmp$breadth_2020["Ideology ID"])`% (2020) and
`r pp(X_cmp$breadth_2024["Ideology ID"])`% (2024), above both the 1984&ndash;2016 value of 27 and the
2008&ndash;2016 value of 32, so the ideological poles are more populated than at any point the author
observed. Every construct became about +0.20 more inter-correlated: ideology's average inter-construct
correlation rose from `r d2(af$avgcorr_8416[4])` to `r d2(X_cmp$avgcorr_2020[4])`, and the *pairwise*
ideology&times;party correlation rose from .44 (original Table 4) to
`r d2(X_t4$table4_2020$Full[X_t4$table4_2020$pair=="IdeoID x Party"])`. Vote-predictiveness rose for
every construct, with partisanship reaching pseudo-R²
`r d2(X_cmp$voteR2_2020["Partisanship"])` (2020) and `r d2(X_cmp$voteR2_2024["Partisanship"])` (2024).
The two value batteries carry item-count artefacts and must be read against their matched baselines,
which the bracketed figures supply. For moral traditionalism the "rise" is almost entirely item-count.
Against its matched 2-item baseline (breadth `r pp(V_F$item_count_decomp$matched_8416[2])`), its
2020/2024 breadth (`r pp(X_cmp$breadth_2020["Moral Tradition"])`/`r pp(X_cmp$breadth_2024["Moral Tradition"])`)
is essentially flat: of the apparent +`r pp(V_F$item_count_decomp$rise_unmatched_2024[2])`-point rise
in 2024, +`r pp(V_F$item_count_decomp$itemcount_pp_2024[2])` is item-count inflation and only
+`r pp(V_F$item_count_decomp$rise_matched_2024[2])` is genuine. Egalitarianism is affected in parallel.
Its matched 4-item baseline is `r pp(V_F$item_count_decomp$matched_8416[1])` rather than the unmatched
`r pp(V_F$item_count_decomp$unmatched_8416[1])`, so of its +`r pp(V_F$item_count_decomp$rise_unmatched_2024[1])`-point
2024 rise, +`r pp(V_F$item_count_decomp$itemcount_pp_2024[1])` is item-count and only
+`r pp(V_F$item_count_decomp$rise_matched_2024[1])` is real. Both constructs' covariance and
correlation rises, which are item-count invariant, survive.

### Where the stratification went

Holding the knowledge measure constant (quiz-versus-quiz) splits the stratification verdict by metric.
**Breadth stratification did not meaningfully collapse; it was always modest on a quiz.** The ideology
breadth ratio was 2.29 on the 1986&ndash;92 quiz, against 4.00 on the interviewer rating, and is
`r d2(X_cmp$breadth_HdivL_2020["Ideology ID"])` in 2020; partisanship was flat then (1.08) and flat now
(`r d2(X_cmp$breadth_HdivL_2020["Partisanship"])`).

**For ideology, policy, and moral traditionalism the quiz-matched vote-potency gradient compressed
sharply; egalitarianism is the exception, where it roughly doubled from a low base** (1.3 on the
1988&ndash;92 quiz to `r d2(X_cmp$voteR2_HdivL_2020["Egalitarianism"])`/`r d2(X_cmp$voteR2_HdivL_2024["Egalitarianism"])`
in 2020/2024). For ideology, the quiz-stratified Highest&divide;Lowest vote-R² ratio was
`r d1(X_cmp$anchor_infogain$vote_R2_HdivL_quiz8892[4])` in 1988&ndash;92, essentially identical to the
interviewer-rated 10.75, with the lowest-knowledge voter group at pseudo-R² .04. In 2020 the same
quiz-based ideology ratio is `r d2(X_cmp$voteR2_HdivL_2020["Ideology ID"])`
(2024: `r d2(X_cmp$voteR2_HdivL_2024["Ideology ID"])`), and the lowest-knowledge voter group's ideology
pseudo-R² has risen to `r d2(x5(X_t5$table5_2020, "Ideology ID", "pseudoR2", "Lower"))` (2020) and
`r d2(x5(X_t5$table5_2024, "Ideology ID", "pseudoR2", "Lower"))` (2024). Low-knowledge voters, whose
ballots barely tracked their ideology in 1988&ndash;92 even on the quiz, now cast substantially
ideologically structured votes. A residual gradient of roughly 2&ndash;2.5&times; remains, so the
paper's strong "meaningful for a knowledgeable minority only" framing no longer holds, though the
pattern is compressed rather than erased.

Because pseudo-R² ratios exaggerate stratification relative to the interpretable probability scale
(RC6c), the same comparison should be checked on average marginal effects (@tbl-rep-ame-ratio). On
the AME scale the ideology gradient was already modest, a Highest&divide;Lower ratio of
`r d2(vb("1988-92 (quiz)","Ideology","AME_H_div_L"))` on the 1988&ndash;92 quiz rather than 10&times;,
and by 2020/2024 it falls *below one* (`r d2(vb("2020 (quiz)","Ideology","AME_H_div_L"))` and
`r d2(vb("2024 (quiz)","Ideology","AME_H_div_L"))`): the low-knowledge group's marginal effect of
ideology on vote probability is now, if anything, slightly *larger* than the high-knowledge group's.
The two metrics thus tell the same directional story, that low-knowledge voters caught up, but they
disagree on magnitude, and the recommended metric shows the vote-potency gradient not merely
compressed but essentially gone. I report the pseudo-R² ratio as the headline for continuity with the
original's own reporting, but flag that its "10.5&times;" starting point is a variance-scale artefact
more than a behavioural fact.

```{r}
#| label: tbl-rep-vote-ratio
#| tbl-cap: "Vote-potency stratification (pseudo-R² Highest÷Lowest), measure-matched. The fair benchmark for the 2020/2024 quiz index is the author's own quiz-stratified 1988–92 table (B5), not the steeper interviewer-rated column. 2020/2024 use merged voter knowledge groups."
ai <- X_cmp$anchor_infogain
vote_ratio <- tibble::tibble(
  Construct = ai$construct,
  `Interviewer 1984–2016` = round(ai$vote_R2_HdivL_8416, 1),
  `Quiz 1988–92 (B5)` = round(ai$vote_R2_HdivL_quiz8892, 1),
  `Quiz 2020` = round(X_cmp$voteR2_HdivL_2020, 2),
  `Quiz 2024` = round(X_cmp$voteR2_HdivL_2024, 2)
)
vote_ratio |>
  gt() |>
  tab_header(title = "Vote-potency gradient: sharply reduced on the quiz-matched test",
             subtitle = "Ideology 10.5× (1988–92 quiz) → ~2.5× (2020) / ~2.0× (2024)") |>
  cols_align("left", columns = Construct)
```

```{r}
#| label: tbl-rep-ame-ratio
#| tbl-cap: "The same Highest÷Lower vote-potency gradient on two metrics: the pseudo-R² ratio (variance scale) and the average-marginal-effect ratio (probability scale). Ideology and party, measure-matched across eras, merged voter knowledge groups. The AME ratio was already ≤ 2× in the original eras and falls below 1 by 2020/2024."
ame_ratio <- tibble::tibble(
  Era = c("1988–92 (quiz)","2020 (quiz)","2024 (quiz)","1984–2016 (interviewer)"),
  `Ideology AME H÷L` = c(vb("1988-92 (quiz)","Ideology","AME_H_div_L"),
                         vb("2020 (quiz)","Ideology","AME_H_div_L"),
                         vb("2024 (quiz)","Ideology","AME_H_div_L"),
                         vb("1984-2016 (interviewer)","Ideology","AME_H_div_L")),
  `Ideology pseudo-R² H÷L` = c(vb("1988-92 (quiz)","Ideology","R2_H_div_L"),
                               vb("2020 (quiz)","Ideology","R2_H_div_L"),
                               vb("2024 (quiz)","Ideology","R2_H_div_L"),
                               vb("1984-2016 (interviewer)","Ideology","R2_H_div_L")),
  `Party AME H÷L` = c(vb("1988-92 (quiz)","Party","AME_H_div_L"),
                      vb("2020 (quiz)","Party","AME_H_div_L"),
                      vb("2024 (quiz)","Party","AME_H_div_L"),
                      vb("1984-2016 (interviewer)","Party","AME_H_div_L")),
  `Party pseudo-R² H÷L` = c(vb("1988-92 (quiz)","Party","R2_H_div_L"),
                            vb("2020 (quiz)","Party","R2_H_div_L"),
                            vb("2024 (quiz)","Party","R2_H_div_L"),
                            vb("1984-2016 (interviewer)","Party","R2_H_div_L"))
)
ame_ratio |>
  gt() |>
  tab_header(title = "On the interpretable scale, ideology's stratification was never 10×",
             subtitle = "AME Highest÷Lower vs pseudo-R² Highest÷Lower") |>
  cols_align("left", columns = Era)
```

### Sorting or constraint?

Every metric above is *bivariate*: breadth, inter-construct correlations, and the ideology vote-R²
all describe ideology on its own. However, the framework this paper replicates warns that measured
"ideology" in mass publics is often party identity in ideological clothing [@kinderkalmoe2017;
@mason2018]. Bivariate ideology potency is exactly what partisan sorting inflates, because if citizens
simply adopt ideological labels that track their existing party, ideology's vote-R² can rise with no
gain in genuine belief-system constraint. The discriminating test is ideology's *incremental*
contribution beyond party identification. For each era and knowledge group, I fit a weighted probit of
Republican two-party vote on party alone and on party + ideology, using identical complete-case rows,
and read off how much ideology adds (@tbl-rep-increment).

```{r}
#| label: tbl-rep-increment
#| tbl-cap: "Ideology's incremental vote contribution net of party ID: the rise in McFadden pseudo-R² from adding ideology to a party-only probit, by era and (merged) voter knowledge group. Measure-matched (interviewer 1984–2016; campaign quiz 1988–92; civics quiz 2020/2024). The final column is the ideology×party correlation, full sample."
incr_wide <- V_A |>
  dplyr::filter(group != "Full") |>
  dplyr::select(era, group, incr_ideo) |>
  tidyr::pivot_wider(names_from = group, values_from = incr_ideo)
full_incr <- V_A$incr_ideo[V_A$group == "Full"]
full_cor  <- V_A$cor_ideo_party[V_A$group == "Full"]
incr_tbl <- tibble::tibble(
  Era = incr_wide$era,
  Full = full_incr,
  Lower = incr_wide$Lower, Middle = incr_wide$Middle, High = incr_wide$High, Highest = incr_wide$Highest,
  `Ideology×party r` = round(full_cor, 2)
)
incr_tbl |>
  gt() |>
  fmt_number(columns = c(Full, Lower, Middle, High, Highest), decimals = 3) |>
  tab_header(title = "Net of party, ideology's independent contribution did not rise",
             subtitle = "Δ pseudo-R² from adding ideology to a party-only vote model") |>
  cols_align("left", columns = Era) |>
  tab_spanner("Increment by knowledge group", columns = c(Lower, Middle, High, Highest))
```

The answer, which is where the review's concern lies, is that it did not. Ideology's increment beyond
party is small in every era, a few pseudo-R² points, and it did not rise overall or at the
low-knowledge floor. The full-sample increment was
`r d2(va("1984-2016 (interviewer)","Full","incr_ideo"))` on the interviewer rating and
`r d2(va("1988-92 (quiz)","Full","incr_ideo"))` on the measure-matched 1988&ndash;92 quiz, against
`r d2(va("2020 (quiz)","Full","incr_ideo"))` in 2020 and `r d2(va("2024 (quiz)","Full","incr_ideo"))`
in 2024, so it is flat and if anything lower. The low-knowledge floor tells the same story: ideology
added `r d2(va("1988-92 (quiz)","Lower","incr_ideo"))` there on the 1988&ndash;92 quiz and adds
`r d2(va("2020 (quiz)","Lower","incr_ideo"))` (2020) and `r d2(va("2024 (quiz)","Lower","incr_ideo"))`
(2024), again no rise. The increments are small and non-monotonic rather than trending. The Middle and
Highest cells do tick up in 2020, to `r d2(va("2020 (quiz)","Middle","incr_ideo"))` and
`r d2(va("2020 (quiz)","Highest","incr_ideo"))`, before 2024's ceiling effect compresses them, since
party alone already reaches pseudo-R² `r d2(va("2024 (quiz)","Full","r2_party"))` there; but nowhere
does adding ideology to party lift fit by more than a handful of points. The asymmetry is the cleaner
summary. Reversing the comparison, adding *party* to an ideology-only model lifts fit by
`r d2(va("2020 (quiz)","Full","incr_party"))` (2020, full sample) against ideology's
`r d2(va("2020 (quiz)","Full","incr_ideo"))`, so party carries the vote while ideology rides along.
Meanwhile the low-knowledge group's *bivariate* ideology vote-R² climbed from about .04 to
`r d2(x5(X_t5$table5_2020, "Ideology ID", "pseudoR2", "Lower"))`/`r d2(x5(X_t5$table5_2024, "Ideology ID", "pseudoR2", "Lower"))`,
and its ideology&times;party correlation climbed from
`r d2(va("1988-92 (quiz)","Lower","cor_ideo_party"))` to
`r d2(va("2020 (quiz)","Lower","cor_ideo_party"))`/`r d2(va("2024 (quiz)","Lower","cor_ideo_party"))`.
Almost the entire rise in low-knowledge ideological vote potency is ideology's growing *alignment with
party*, not a growing independent contribution.

The honest reading is descriptive rather than adjudicating: net of party, ideology's independent
electoral contribution changed little, so the striking bivariate rise reflects ideology and party
moving together. This does not prove "sorting, not constraint", since two genuinely co-evolving
constructs would leave the same signature, and I cannot separate citizens adopting party-aligned
ideological labels from parties becoming more ideologically defined. Nor does it fully rescue the
innocence thesis. On the interpretable probability scale a modest independent ideology effect persists
in every cell: the joint-model ideology AME runs from about
`r d2(va("2020 (quiz)","Lower","ame_ideo_joint"))` at the low-knowledge floor to
`r d2(va("2020 (quiz)","Highest","ame_ideo_joint"))` at the top in 2020, modest and present
everywhere, with a mild residual gradient. The coherence and stability gradients below are also not
bivariate-vote artefacts. It does, however, qualify the strong "time-bound" claim, because the part of
the apparent shift that is genuine *independent* ideological constraint is small and did not grow. What
rose most is how tightly ideology and party now travel together.

### Stability

On the `r format(V_C$n_panel, big.mark=",")`-case 2016&ndash;2020&ndash;2024 panel, squared
2020&rarr;2024 continuity correlations rose for every construct relative to the original 1992&ndash;96
panel, which spans the same four-year gap (@tbl-rep-stability). Ideological identification rose by
roughly three-quarters, from .37 to
`r d2(X_stab$stability_2020_2024$Full[X_stab$stability_2020_2024$construct=="Ideology ID"])`
(&times;`r d2(X_stab$stability_2020_2024$Full[X_stab$stability_2020_2024$construct=="Ideology ID"]/0.37)`).
Even the low-knowledge group's ideology stability
(`r d2(X_stab$stability_2020_2024$Lower[X_stab$stability_2020_2024$construct=="Ideology ID"])`) exceeds
the original *full-sample* value. Because the reduced value batteries add noise that attenuates
test&ndash;retest correlations, the egalitarianism and moral-traditionalism rises are if anything
conservative; that attenuation argument, however, is not the whole story on the direction of bias.
These estimates come from respondents in a *three-wave* panel (2016&ndash;2020&ndash;2024), and by the
third interview panel conditioning, in which respondents crystallise or anchor to remembered prior
answers, can *inflate* test&ndash;retest correlations beyond a fresh two-wave design. Measurement noise
and conditioning push in opposite directions, and which dominates the net bias is not determinable
here, so the stability rise should be read as directional rather than as a precise magnitude.

**Panel attrition.** Multi-wave panels retain the politically engaged selectively, so the panel could
over-represent exactly the respondents whose attitudes are most stable, and could do so differentially
by knowledge. Retention of the 2020 cross-section into the 2024 panel wave is indeed mildly
knowledge-graded: `r d1(V_C$composition_5grp$retention_pct[V_C$composition_5grp$group=="Lowest"])`% of
the lowest quiz group was retained against
`r d1(V_C$composition_5grp$retention_pct[V_C$composition_5grp$group=="Highest"])`% of the highest
(overall `r d1(V_C$overall_retention_pct)`%), so the lowest group is slightly under-represented in the
panel, at `r d1(V_C$composition_5grp$panel_pct[V_C$composition_5grp$group=="Lowest"])`% of panel cases
against `r d1(V_C$composition_5grp$crosssection_pct[V_C$composition_5grp$group=="Lowest"])`% of the
cross-section. The bias this induces is small, however. Post-stratifying the panel back to the 2020
cross-sectional quiz distribution barely moves the stability estimates, taking reweighted ideology
stability to `r d2(V_C$stability_reweighted$poststrat_5grp[V_C$stability_reweighted$construct=="Ideology ID"])`
against the unweighted `r d2(V_C$stability_reweighted$unweighted[V_C$stability_reweighted$construct=="Ideology ID"])`,
and applying the ANES panel weight (`V240106b`) lowers it only to
`r d2(V_C$stability_reweighted$panelwt_V240106b[V_C$stability_reweighted$construct=="Ideology ID"])`
(@tbl-rep-stab-reweight). Even under the panel weight, low-knowledge ideology stability stays well
above the original full-sample value, so differential attrition does not manufacture the result,
though it likely inflates the level modestly.

```{r}
#| label: tbl-rep-stab-reweight
#| tbl-cap: "Attrition sensitivity for the 2020→2024 stability estimates. Unweighted (primary, matching the original's unweighted Table 3); post-stratified to the 2020 cross-sectional quiz-knowledge distribution over the five quiz groups; and multiplied by the ANES V240106b panel weight."
V_C$stability_reweighted |>
  gt() |>
  cols_label(construct = "Construct", unweighted = "Unweighted", poststrat_5grp = "Post-strat (5 groups)",
             panelwt_V240106b = "× panel weight", poststrat_x_panelwt = "Both") |>
  tab_header(title = "Reweighting for differential attrition barely moves stability",
             subtitle = "Squared 2020→2024 continuity correlations, full sample") |>
  cols_align("left", columns = construct)
```

```{r}
#| label: tbl-rep-stability
#| tbl-cap: "Squared continuity (test–retest) correlations, 2020→2024, versus the original 1992–96 panel, by 2020-wave quiz knowledge."
st <- X_stab$stability_2020_2024
st_orig <- X_stab$original_1992_96
stab_tbl <- tibble::tibble(
  Construct = st$construct,
  `1992–96 (orig)` = st_orig$r2_1992_96[match(st$construct, st_orig$construct)],
  `2020→2024 Full` = st$Full,
  Lower = st$Lower, Middle = st$Middle, High = st$High, Highest = st$Highest
)
stab_tbl |>
  gt() |>
  tab_header(title = "Stability rose for every construct",
             subtitle = "Ideological identification rose ×1.73 (.37 → .64)") |>
  cols_align("left", columns = Construct) |>
  tab_source_note(gt::md(sprintf(
    "Merged voter knowledge groups; among the %s panel cases the 2020-quiz group shares are Lower %s%%, Middle %s%%, High %s%%, Highest %s%%.",
    format(V_C$n_panel, big.mark=","),
    d1(V_C$composition_4grp$panel_pct[1]), d1(V_C$composition_4grp$panel_pct[2]),
    d1(V_C$composition_4grp$panel_pct[3]), d1(V_C$composition_4grp$panel_pct[4]))))
```

### The shrinking middle

The ideological middle contracted. The Moderate/HTMA bar fell from about 49% in the original era
(50.0% reproduced) to `r d1(X_fig1$y2020$ideo_moderate_or_htma)`% in 2020 and
`r d1(X_fig1$y2024$ideo_moderate_or_htma)`% in 2024, as the liberal and conservative categories grew
(@fig-ideo-eras). "Haven't thought much about it" persists on the web at
`r d1(X_fig1$y2020$pct_htma)`% (2020) / `r d1(X_fig1$y2024$pct_htma)`% (2024), and opinionation still
tracks knowledge: mean quiz knowledge is
`r d2(X_opin$y2020$summary$value[X_opin$y2020$summary$stat=="knowledge_HTMA"])` for HTMA identifiers
versus `r d2(X_opin$y2020$summary$value[X_opin$y2020$summary$stat=="knowledge_libcon"])` for those who
place themselves.

Because RC2 showed the breadth statistic is partly a nonresponse statistic, the breadth rise itself
must be decomposed into genuine polar movement and reduced nonattitude. All figures here are weighted
on a common basis, so the accounting is internally consistent (the earlier unweighted web-HTMA figures
differ slightly by design). Weighted polar-half breadth rose from
`r d1(V_D$ideology["orig","base"])`% to `r d1(V_D$ideology["y2020","base"])`% (2020) /
`r d1(V_D$ideology["y2024","base"])`% (2024); among *attitude-holders only* (dropping HTMA) it rose
from `r d1(V_D$ideology["orig","holders"])`% to `r d1(V_D$ideology["y2020","holders"])`% /
`r d1(V_D$ideology["y2024","holders"])`%, while the weighted HTMA share fell from
`r d1(V_D$ideology["orig","htma"])`% to `r d1(V_D$ideology["y2020","htma"])`% /
`r d1(V_D$ideology["y2024","htma"])`%. Decomposing the +`r d1(V_D$decomp_2020["total_rise"])`-point
2020 rise, only +`r d1(V_D$decomp_2020["nonresponse_component"])` points come from reduced nonresponse
and +`r d1(V_D$decomp_2020["polar_component"])` from genuine polar movement among attitude-holders
(2024: +`r d1(V_D$decomp_2024["nonresponse_component"])` nonresponse, +`r d1(V_D$decomp_2024["polar_component"])`
polar). The breadth rise is therefore mostly real polarisation rather than people simply abandoning
"haven't thought much", though the nonresponse component is non-trivial and, per RC2, the *original*
gradient was more nonresponse-driven than the level rise is.

Whether the contraction is a web-mode effect is harder to settle, because 2020 is
`r d1(V_E$mode_table$n_administered[V_E$mode_table$mode=="Web"]/sum(V_E$mode_table$n_administered)*100)`%
web and the face-to-face-successor modes are thin (@tbl-rep-mode). The HTMA *component* is not
mode-driven: HTMA runs `r d1(V_E$mode_table$htma_pct[V_E$mode_table$mode=="Web"])`% on the web,
`r d1(V_E$mode_table$htma_pct[V_E$mode_table$mode=="Video"])`% on video, and
`r d1(V_E$mode_table$htma_pct[V_E$mode_table$mode=="Phone"])`% on phone. The *Moderate*
self-placement rate, about which HTMA flatness says nothing, is not obviously flat, however, running
`r d1(V_E$mode_table$moderate_pct[V_E$mode_table$mode=="Web"])`% on the web against
`r d1(V_E$mode_table$moderate_pct[V_E$mode_table$mode=="Phone"])`% on phone. With only
`r V_E$mode_table$n_administered[V_E$mode_table$mode=="Video"]` video and
`r V_E$mode_table$n_administered[V_E$mode_table$mode=="Phone"]` phone respondents, the non-web polar
share is estimated far too imprecisely (web polar
`r d1(V_E$mode_table$polar_overall[V_E$mode_table$mode=="Web"])`%, 95% CI
`r d1(V_E$mode_table$polar_overall_lo[V_E$mode_table$mode=="Phone"])`&ndash;`r d1(V_E$mode_table$polar_overall_hi[V_E$mode_table$mode=="Phone"])`
on phone) to rule a mode effect in or out. The honest statement is therefore a narrow one: the HTMA
share is mode-independent, but whether the moderate-to-polar shift is mode-driven cannot be determined
from the tiny non-web subsamples. Ideology's bivariate vote pseudo-R² is close across the web/non-web
split (`r d2(V_E$ideo_voteR2_web$r2[V_E$ideo_voteR2_web$group=="Web"])` web against
`r d2(V_E$ideo_voteR2_web$r2[V_E$ideo_voteR2_web$group=="Non-web"])` non-web), which is reassuring but
rests on only `r V_E$ideo_voteR2_web$n[V_E$ideo_voteR2_web$group=="Non-web"]` non-web voters.

```{r}
#| label: tbl-rep-mode
#| tbl-cap: "Ideology metrics by 2020 interview mode. Moderate and HTMA are shown separately (the composite is what shrank). Non-web cells are tiny — n's are given and the polar share carries a wide interval — so mode effects cannot be pinned down. Percentages among administered cases; correlation among self-placers."
V_E$mode_table |>
  dplyr::transmute(Mode = mode, `n (admin.)` = n_administered,
                   `Polar %` = polar_overall, `Moderate %` = moderate_pct, `HTMA %` = htma_pct,
                   `Mean |ideo|` = mean_abs_ideo, `Ideo×party r` = cor_ideo_party) |>
  gt() |>
  tab_header(title = "2020 ideology by mode: HTMA is flat, the rest is too noisy to call",
             subtitle = "Web n ≈ 7,800; video ≈ 360; phone ≈ 140") |>
  cols_align("left", columns = Mode)
```

```{r}
#| label: fig-ideo-eras
#| fig-cap: "The ideological middle shrank as the poles grew. Seven-category liberal–conservative self-placement across three eras; the Moderate/HTMA category fell from 50% to about 35%. 1984–2016 is the reproduced ANES distribution; 2020/2024 are the replication distributions."
#| fig-height: 6
eras <- bind_rows(
  R_fig1$ideology     |> transmute(cat = ideo_cat, label, pct, era = "1984–2016"),
  X_fig1$y2020$ideology |> transmute(cat = ideo_cat, label, pct, era = "2020"),
  X_fig1$y2024$ideology |> transmute(cat = ideo_cat, label, pct, era = "2024")
) |>
  mutate(label = factor(label, levels = R_fig1$ideology$label),
         era = factor(era, levels = c("1984–2016","2020","2024")),
         is_mid = label == "Moderate/HTMA")

ggplot(eras, aes(x = pct, y = forcats::fct_rev(label))) +
  geom_col(aes(alpha = is_mid), fill = "#2f6b8f", width = 0.72) +
  geom_text(aes(label = sprintf("%.0f", pct)), hjust = -0.15, size = 3, colour = "grey25") +
  scale_alpha_manual(values = c(`TRUE` = 1, `FALSE` = 0.55), guide = "none") +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15))) +
  facet_wrap(~era) +
  labs(x = "% of respondents", y = NULL,
       title = "The ideological middle emptied out",
       subtitle = "Moderate/HTMA (darker) fell from 50% to ~35% as the poles filled") +
  theme_minimal(base_size = 12) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_blank(),
        plot.title = element_text(face = "bold"),
        strip.text = element_text(face = "bold"))
```

<details>
<summary>Full replication grids (correlations, reliability, vote, by knowledge group)</summary>

```{r}
#| label: tbl-rep-t4-full
#| tbl-cap: "Replication Table 4 analogue (2020): construct correlations by quiz-knowledge group."
X_t4$table4_2020 |>
  mutate(across(Full:Highest, ~round(.x,2))) |> select(-N) |>
  gt() |> tab_header(title = "2020 correlations by knowledge group")
```

```{r}
#| label: tbl-rep-t5-full
#| tbl-cap: "Replication Table 5 analogue (2020): vote probit by merged voter knowledge group."
X_t5$table5_2020 |>
  mutate(across(Full:Highest, ~round(.x,2))) |> select(-H_div_L_R2) |>
  gt() |> tab_header(title = "2020 vote probit coefficient / SE / pseudo-R²")
```

</details>

### Verdicts

Set against the four original headline claims, the picture is consistent across metrics
(@tbl-rep-verdicts): the descriptive portrait shifted while the underlying knowledge mechanism
persisted in weakened form.

```{r}
#| label: tbl-rep-verdicts
#| tbl-cap: "The four original headline claims in 2020/2024."
verd <- tibble::tibble(
  `Original claim` = c(
    "1. Polar/coherent/potent ideology only for the knowledgeable ~20–30%",
    "2. Partisanship dominates ideology on every metric",
    "3. Knowledge stratification is enormous (info-gain 1.6–11×)",
    "4. Values carry the same knowledge-dependent limits as ideology"),
  `Verdict in 2020/2024` = c(
    "Not replicated as stated. The knowledge gradient persists, but the low-knowledge floor rose enough that ideology is no longer confined to a minority, though net of party the independent rise is small.",
    "Replicated, though the gap narrowed.",
    "Splits by metric. Breadth is roughly unchanged, having always been modest on a quiz, while vote potency is sharply reduced but not eliminated.",
    "Replicated structurally. Values track ideology: all rise, and all flatten."),
  `Key evidence` = c(
    "Lowest-K ideology vote-R² .04→.25/.31; ideology×party corr lowest group .06→.59",
    "Party÷ideology vote-R² 2.1×→1.4×; ideology stability .37→.64 nears party's .78",
    "Vote-R² H÷L quiz ideology 10.5→~2.5/2.0; breadth H÷L quiz ideology 2.29→1.9",
    "Avg corr egal .36→.54, MT .31→.53, parallel to ideology .39→.64")
)
verd |>
  gt() |>
  tab_header(title = "Two-sided result: the portrait is time-bound, the mechanism partly persists") |>
  cols_align("left", columns = everything())
```

## GSS Conceptual Replication

The GSS test asks whether the pattern recurs in a *different survey* with a *different knowledge proxy*
(WORDSUM vocabulary), addressing the nomination's cross-dataset priority. The result is an instructive
**depth-versus-breadth dissociation**. The depth gradients replicate almost identically. The
ideology&times;party correlation rises from `r d2(G3$wordsum$Lowest[G3$wordsum$pair=="Ideology x Party"])`
in the lowest WORDSUM group to `r d2(G3$wordsum$Highest[G3$wordsum$pair=="Ideology x Party"])` in the
highest (full-sample `r d2(G3$wordsum$Full[G3$wordsum$pair=="Ideology x Party"])`), tracking Kalmoe's
ANES interviewer-rating gradient of .06&rarr;.68 almost exactly. Policy-index reliability rises
`r d2(G2$wordsum$Lowest[G2$wordsum$stat=="alpha"])`&rarr;`r d2(G2$wordsum$Highest[G2$wordsum$stat=="alpha"])`,
and ideology's vote pseudo-R² is about 15&times; larger at the top
(`r d2(G4$Highest[G4$predictor=="Ideology" & G4$stat=="pseudoR2"])`) than the bottom
(`r d2(G4$Lowest[G4$predictor=="Ideology" & G4$stat=="pseudoR2"])`). Partisanship dominates ideology in
*every* knowledge group with a comfortable margin: its vote pseudo-R² runs
`r d2(G4$Lowest[G4$predictor=="Partisanship" & G4$stat=="pseudoR2"])` (lowest) to
`r d2(G4$Highest[G4$predictor=="Partisanship" & G4$stat=="pseudoR2"])` (highest), against ideology's
`r d2(G4$Lowest[G4$predictor=="Ideology" & G4$stat=="pseudoR2"])` and
`r d2(G4$Highest[G4$predictor=="Ideology" & G4$stat=="pseudoR2"])`. Even in the top group, where
ideology is strongest, partisanship leads by
`r d2(G4$Highest[G4$predictor=="Partisanship" & G4$stat=="pseudoR2"] - G4$Highest[G4$predictor=="Ideology" & G4$stat=="pseudoR2"])`.
This "every group" claim is a within-group comparison, not the full-sample partisanship value set
against the top-group ideology value.

The **breadth** gradient, by contrast, does not replicate: the ideology polar-half ratio is only
`r d2(G1$wordsum$H_div_L[G1$wordsum$construct=="Ideology"])` with WORDSUM, and 1.40 with education,
against Kalmoe's 4.0. The explanation is not a weak moderator, because the *same* WORDSUM bins produce
a clean 7&times; correlation gradient and a 15&times; vote-potency gradient, so verbal ability tracks
the depth metrics faithfully. What differs is the metric. GSS POLVIEWS offers no "haven't thought
much" escape, and so forces a substantive 7-point placement, with only `r d1(G6$full*100)`% declining
to place. Since RC2 showed the ANES breadth gradient is largely a nonresponse gradient, removing the
nonresponse channel is exactly what should flatten it: low-knowledge respondents choose polar *labels*
almost as often as sophisticates once they cannot opt out, but only among the knowledgeable do those
labels cohere with party, policy, and the vote. Because breadth is primarily a nonresponse artefact,
the strength of the knowledge proxy is beside the point for it; the proxy's strength is instead what
lets it reproduce the depth gradients, which it does. The deeper and more consequential half of the
thesis replicates cleanly across the dataset boundary.

The GSS also shows the two-sided temporal story in one dataset. Full-sample ideology&times;party
constraint tripled from the 1970s (.21) to the 2020s (.64), the well-documented partisan sort
[@levendusky2009; @mason2018], yet the knowledge gap persists in every decade: even in the 2020s the
bottom-knowledge correlation (.30) sits far below the top's (.77). Sorting raised the average without
erasing stratification, so an unstratified 2020s snapshot still overstates ideological constraint for
most citizens.

## Discussion

**What the original got right: nearly everything, reproducibly so.** The reproduction regenerates the
published tables cell by cell from the same public data, and every substantive conclusion survives a
demanding robustness suite that validates against the author's own supplement. The three anomalies are
documentation-level: a probable N transposition, one unreproducible panel cell, and one undocumented
panel item set. For a secondary-data paper with no code package, this is an unusually clean bill of
health, and it should raise rather than lower confidence in the original analysis.

**What is time-bound, and how much of it is sorting.** On the author's own tests, applied to data he
never saw, the 2020s public looks markedly more ideological than the public he described. Coherence,
polarity, stability, and bivariate vote potency all rose, and on those bivariate metrics the
low-knowledge floor rose furthest, so the strong claim that meaningful ideology is confined to a
knowledgeable minority does not hold as stated in 2020/2024. However, the party-controlled test
qualifies what that rise means. Net of party identification, ideology's *independent* contribution to
the vote did not grow overall (full-sample increment .03&ndash;.04 across every era) or at the
low-knowledge floor, and the dramatic rise in low-knowledge ideological voting is almost entirely
ideology's growing alignment with party rather than a new stratum of independent ideological
constraint. The honest description is therefore that ideology and party became far more tightly
aligned, especially where they used to diverge, rather than that everyone simply became more
ideological. The gradient did compress, with the top-to-bottom vote-potency ratio falling from roughly
10&times; to about 2&times; on pseudo-R² and, on the interpretable AME scale, below one. This is a
change in the *shape* of the gradient, concentrated at the low end, rather than a uniform level shift:
ideology rose at every knowledge level while the gains were largest at the bottom. The mechanism did
not vanish either. Coherence remains stratified, a modest independent ideology effect persists at every
level, and partisanship still leads every metric, consistent with the "party over ideology" literature
[@barberpope2019; @mason2018].

**Implications for the four best practices.** All four survive. Stratifying by knowledge still matters,
since the gradient is compressed rather than gone, and an unstratified estimate still misleads.
Convenience samples still over-represent the knowledgeable [@clifford2015]. Distinguishing ideology from
partisanship matters more than ever: the two are now highly correlated, partisanship still leads on
every metric, and conflating them risks attributing to ideology what party identity produces, as the
party-controlled analysis above shows concretely. And full-sample estimates remain
knowledge-conditioned. To these I would add three refinements that this project surfaces: **report how
nonattitudes are handled**, since the breadth statistic is partly a nonresponse statistic (RC2, GSS);
**prefer marginal effects to pseudo-R² ratios** when describing how strongly a predictor stratifies,
since the two can differ by an order of magnitude (RC6c); and **beware that a knowledge quiz and an
interviewer rating are not interchangeable stratifiers**, since the quiz yields systematically shallower
gradients, so cross-era comparisons must be measure-matched.

**Limitations.** Several constraints bound these conclusions. The replication's knowledge measure is a
quiz rather than the interviewer rating (handled by the quiz-versus-quiz benchmark, though a residual
instrument difference remains); 2020 and 2024 are single elections compared against multi-year pools
(which can inflate the apparent jump, though the 2008&ndash;2016 pool bounds this); the mode shifted to
web without an interviewer to probe nonresponse (its effect on HTMA/DK is modest but not separable from
the era); the value batteries are reduced (handled via matched items and covariance); the analysis is
single-country except for the GSS; and two panel rows are reconstructions rather than exact
reproductions. The headline ratios are also imprecise, since they divide two pseudo-R² values estimated
in subsamples. A nonparametric bootstrap of the 2020 ideology Highest&divide;Lower vote-potency ratio
(`r V_G$B` draws; Lower $n = `r format(V_G$n_lower, big.mark=",")`$, Highest
$n = `r format(V_G$n_highest, big.mark=",")`$) puts the point estimate `r d2(V_G$point)` in a 95%
interval of `r d2(V_G$ci[1])`&ndash;`r d2(V_G$ci[2])`, well below the original era's ~10&times; but too
wide to distinguish the 2020 ratio sharply from the 2024 one. The ratios should therefore be read as
order-of-magnitude compressions rather than precise point estimates. None of these undoes the direction
or rough magnitude of the results, all of which were independently verified.

## Reproducibility Statement

All analyses run in R 4.5.3 [@rcoreteam] with `dplyr`, `tidyr`, `psych` [@rpsych], `sandwich`
[@rsandwich], `MASS`, and `marginaleffects`; tables and figures here use `gt` and `ggplot2`. Every
pipeline is deterministic (no randomness) and was verified to regenerate byte-identically. The full chain
runs from the project root:

```
Rscript R/01_prepare_cdf.R        # ANES CDF → cdf_analysis.rds
Rscript R/02_reproduce_main.R     # Tables 1,2,4,5, Fig 1, footnote 5, in-text
Rscript R/03_robustness.R         # RC1–RC8
Rscript R/04_panels.R             # Table 3 (panels)
Rscript R/05_gss.R                # GSS conceptual replication
Rscript R/06_replication_anes.R   # ANES 2020/2024 + panel
Rscript R/07_revision.R           # post-review analyses → output/revision/*.rds
# Independent verification pass (§2.4): hand-rolled recomputation of every headline
Rscript output/verification/v_reproduction.R
Rscript output/verification/v_robustness.R
Rscript output/verification/v_panels.R
Rscript output/verification/v_replication.R
Rscript output/verification/v_gss.R
quarto render manuscript/manuscript.qmd
```

Data sources: ANES Cumulative Data File (1948&ndash;2016 vintage via the `anesr` mirror; 1948&ndash;2024
subset via SDA Berkeley, accessed 2026-07-03), the ANES 1990&ndash;92 / 1992&ndash;96 / 2000&ndash;02
panels, the 2024 knowledge battery and 2016&ndash;20&ndash;24 panel link (SDA), and the GSS 1972&ndash;2024
Cumulative Datafile (SDA). This manuscript reads only the precomputed `.rds` outputs, so it renders quickly
and independently of the analysis scripts. Repository layout: `R/` (numbered pipeline), `data/raw/`
(downloaded subsets), `output/` (per-module `.rds`, `comparison.csv`, and reports), `notes/` (decisions log
and data provenance), `manuscript/` (this document and `references.bib`). No original code or data package
exists for Kalmoe (2020); the pipeline reconstructs the analysis from the paper and its author-hosted online
supplement.

## Workflow and AI-Assistance Statement {.unnumbered}

The empirical work reported here was executed by an AI-agent workflow (Anthropic's Claude, via Claude Code)
under the direction and review of the human author. The agents located and downloaded the public data,
reconstructed the analysis pipeline, adjudicated unstated methodological choices against the anchors the
original tables supply, ran an independent adversarial verification of every headline number (which found
and corrected one bug, reported in §3&ndash;4), and drafted this manuscript from the resulting outputs. The
human author defined the scope and framing, made the consequential judgment calls, and is responsible for
the content. Every quantitative claim in the text is loaded programmatically from, or checked against, the
precomputed analysis outputs rather than written from memory. This disclosure is offered in the spirit of
the transparency norms the FORRT initiative promotes [@forrt_nomination; @freese2017].

This manuscript was subsequently revised in response to an automated peer review generated by a
"coarse" review pipeline (a structured, multi-stage critique run with a local model). The full review is
archived at `coarse-output/manuscript_review.md`. In response, six new analyses were added &mdash; the
party-controlled ("sorting versus constraint") vote models, AME-based stratification ratios, panel
attrition and reweighting, breadth with and without nonattitudes, a within-2020 mode comparison, and a
set of smaller checks &mdash; all in `R/07_revision.R` (outputs in `output/revision/`), and the
interpretive claims were tightened accordingly. The revision changed one headline emphasis on the
evidence: the party-controlled test showed that ideology's independent vote contribution did not rise,
so the "time-bound" reading is now qualified as substantially an ideology&ndash;party alignment shift.

## References {.unnumbered}
